---
title: "Parallelizing *tidyverse* workflows"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[multidplyr](https://github.com/hadley/multidplyr) allows to split the analisyis of evenly-sized subsamples of large datasets (~10m obs) by dispatching across multiple cores. See the *multidplyr* [vignette](https://github.com/hadley/multidplyr/blob/master/vignettes/multidplyr.md) for details and examples.

## Installation

*multidplyr* is a library under development: installation requires the [devtools](https://cran.r-project.org/web/packages/devtools/index.html) library to compile from source – and [rtools](https://cran.r-project.org/bin/windows/Rtools/) on Windows.

```{R message=FALSE, warning=FALSE}
devtools::install_github("hadley/multidplyr")
```


## Example: country-level regression

From cross-country firm-level data, we want to fit country-wise regression models of average wages on firm controls.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(multidplyr)
```


Create dummy data:


```{r}
countries <- c("Italy", "France", "Germany", "Spain", "UK") %>% as.factor()

# Number of observations for each country
n <- 1000000 

firm_data <- tibble(
  country = rep(countries, each = n),
  profits = runif(n = n*length(countries), min = 15, max = 50),
  unionized = sample(c(TRUE, FALSE), size = n*length(countries), replace = TRUE),
  share_tertiary = runif(n*length(countries)),
  employees = runif(n = n*length(countries), min = 1, max = 300) %>% round()
)

firm_data
```

Initialize a cluster and split the data by country into “shards” using `partition()`. Similar to `group_by()`, but will send each group to a different cluster.

```{r}
firm_data_part <- partition(firm_data, country)
```

Then, fit a model by groups using a `do()` expression:
```{r}
firm_data_part %>% 
  do(model_wage = lm(profits ~ employees + unionized +employees + share_tertiary,  data = .))
```
